---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE, message = FALSE
)
```

# chdp

<!-- badges: start -->
<!-- badges: end -->

The chdp package is developed for implementing covariate-dependent hierarchical Dirichlet process, which allows for clustering across related datasets and incorporating the information of external covariate. The covariate can be flexibly included through different kernel functions. Two different applications are included in the package, one for single-cell clustering with a Gaussian kernel, and the other for clustering time-series data based on vector autoregression (VAR) with a periodic kernel.

## Installation

You can install the development version of chdp from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("huizizhang949/chdp")
```

# Example - clustering single-cell RNA-sequencing data

First, prepare the datasets including the external covaraite - latent time:

```{r, data-gkernel}
library(chdp)
# G - dimension (number of genes)
# C - number of data points in each dataset
# J - number of clusters
# Y - matrix: rows are observations
# z - data allocation
# t - covariate (latent time)
# p_j - covariate-dependent time probabilities
G=100; C=c(200,300); J=3
# first dataset
Y1 <- sim1.data[1:C[1],1:G]; z1 <- sim1.data[1:C[1],G+1]; 
t1 <- sim1.data[1:C[1],G+2]; p_j1 <- sim1.data[1:C[1],-(1:(G+2))];
# second dataset
Y2 <- sim1.data[(C[1]+1):sum(C),1:G]; z2 <- sim1.data[(C[1]+1):sum(C),G+1]; 
t2 <- sim1.data[(C[1]+1):sum(C),G+2]; p_j2 <- sim1.data[(C[1]+1):sum(C),-(1:(G+2))]
```

## Consensus clustering

To implement consensus clustering, first choose the width (number of chains) and depth (length of each chain). Below is an example of running consensus clustering, with 100 chains and 200 iterations on 8 cores
```{r, Width-Depth}
Width <- 100
Depth <- 200
```

```{r, run-consensus, eval=FALSE}
# for reproducibility
set.seed(14663)
seeds <- sample(1:1e8, Width)
# parallel computing on 8 cores
consensus_result <- pblapply(1:Width, function(i) {

  set.seed(seeds[i])
  b_inits <- runif(2,-5,5)
  alpha_inits <- runif(1,0,10)
  alpha_0_inits <- runif(1,0,10)

  gkernelHDP_mcmc(Y = list(t(Y1), t(Y2)), t = list(t1, t2), J = 7, niter = Depth, burn_in = 0, thinning = 1,
                 empirical = TRUE, empirical_z = TRUE,
                 b_initial = b_inits, alpha_initial = alpha_inits, alpha_0_initial = alpha_0_inits,
                 beta.mean = 0.6)
}, cl=8)
```

Specify a range of condidate values (the first element is the baseline):

```{r}
Ds <- c(1,seq(20,200,by=10))
Ws <- c(1,seq(10,100,by=5))
```

Compute posterior similarity matrices (PSM) for different combination W and D

```{r, eval=FALSE}
psm_list <- psm_list_consensus(Ws = Ws, Ds = Ds, consensus_result = consensus_result)
```

Choose a suitable value based on mean absolute difference between PSMs
```{r, eval=FALSE}
plot_consensus(Ws = Ws, Ds = Ds, psm_list = psm_list)
```

```{r, out.width="0.3\\linewidth", include=TRUE, fig.align="center", echo=FALSE}
knitr::include_graphics("demo/figures/plot_consensus.pdf")
```



